\documentclass{extarticle}
\sloppy
\input{packages.tex}
\input{math_commands.tex}


\title{\vspace{-2em}Chapter 4: Polynomials}
\author{\emph{Linear Algebra Done Right}, by Sheldon Axler}
\date{}

\begin{document}
\maketitle 

\begin{definition}[Re z, Im z]
    Suppose \(z = a + b i\), where \(a\) and \(b\) are real numbers. 
    \begin{itemize}
        \item The \textbf{real part} of \(z\), denoted by Re\(z\), is defined by Re\(z=a\).
        \item The \textbf{imaginary part} of \(z\), denoted by Im\(z\), is defined by Im\(z=a\).
    \end{itemize}
\end{definition}


\begin{definition}[complex conjugate, \(\zo\), absolute value, \(|z|\)]
    Suppose \(z \in \C\). 
    \begin{itemize}
        \item The \textbf{complex conjugate} of \(z \in \C\), denoted by \(\zo\), is defined by 
        \[\zo = \text{Re}z - \text{Im}z\]
        \item The \textbf{absolute value} of a complex number \(z\), denoted by \(|z|\), is defined by 
        \[|z| = \sqrt{(\text{Re}(z))^2 + (\text{Im}(z))^2}\]
    \end{itemize}
\end{definition}

A function \(p \colon \F \to \F\) is called a polynomial of degree \(m\) if there exists 
\(a_0, \ldots, a_m \in \F\) with \(a_m \neq 0\) such that 
\[p(z) = a_0 + a_1z + \cdots + a_m z^m\]
for all \(z \in \F\). 

\begin{definition}[zero of a polynomial]
    A number \(\lambda \in \F\) is called a \textbf{zero} (or \textbf{root}) of a polynomial 
    \(p \in\Pc(\F)\) if 
    \[p(\lambda) = 0.\]
\end{definition}


\begin{lemma}[each zero of a polynomial corresponds to a degree-one factor]
    Suppose \(m\) is a positive integer and \(p \in \Pc(\F)\) is a polynomial of degree \(m\). 
    Suppose \(\lambda \in \F\). Then \(p(\lambda) = 0\) if and only if there exists a polynomial 
    \(q \in \Pc(\F)\) of degree \(m - 1\) such that 
    \[p(z) = (z - \lambda)q(z)\]
    for every \(z \in \F\). 
\end{lemma}

\begin{thm}[degree \(m\) implies at most \(m\) zeros]
    Suppose \(m\) is a positive integer and \(p \in \Pc(\F)\) is a polynomial of degree \(m\). 
    Then \(p\) has at most \(m\) zeros in \(\F\). 
\end{thm}

\begin{thm}[division algorithm for polynomials]
    Suppose that \(p, s \in \Pc(\F)\) with \(s \neq 0\). Then there exist unique polynomials 
    \(q, r \in \Pc(\F)\) such that 
    \[p = sq + r\]
    and \(\deg r < \deg s\). 
\end{thm}

\begin{thm}[fundamental theorem of algebra, first version]
    Every nonconstant polynomial with complex coefficients has a zero in \(\C\). 
\end{thm}

\begin{thm}[fundamental theorem of algebra, second version]
    If \(p \in \Pc(\C)\) is a nonconstant polynomial, then \(p\) has a unique factorization 
    (except for the order of the factors) of the form 
    \[p(z) = c(z - \lambda_1) \cdots (z - \lambda_m),\]
    where \(c, \lambda_1, \ldots, \lambda_m \in \C\). 
\end{thm}

\begin{lemma}[polynomials with real coefficients have nonreal zeros in pairs]
    Suppose \(p \in \Pc(\C)\) is a polynomial with real coefficients. If \(\lambda \in \C\)
    is a zero of \(p\), then so is \(\overline{\lambda}\). 
\end{lemma}

\begin{thm}[factorization of a quadratic polynomial]
    Suppose \(b, c \in \R\). Then there is a polynomial factorization of the form 
    \[x^2 + bx + c = (x - \lambda_1)(x - \lambda_2)\]
    with \(\lambda_1, \lambda_2 \in \R\) if and only if \(b^2 \geq 4c\). 
\end{thm}


\begin{thm}[factorization of a polynomial over \(\R\)]
    Suppose \(p \in \Pc(\R)\) is a nonconstant polynomial. Then \(p\) has a unique factorization 
    (except for the order of the factors) of the form 
    \[p(x) = c(x-\lambda_1)\cdots(x - \lambda_m)(x^2 + b_1x + c_1) \cdots(x^2 + b_M x + c_M)\]
    where \(c, \lambda_1, \ldots, \lambda_m, c_1, \ldots, c_M \in \R\), with 
    \(b_k^2 < 4 c_k\) for each \(k\). 
\end{thm}




\newpage 

\begin{problem}{1}
    Suppose \(w, z \in \C\). Verify the following equalities and inequalities. 
    \begin{enumerate}[label=(\alph*)]
        \item \(z + \zo = 2 \Re z\) 
        \item \(z - \zo = 2 (\Im z)i\)  
        \item \(z \zo = |z|^2\) 
        \item \(\overline{w + z} = \wo + \zo\) and \(\overline{wz} = \wo \zo\) 
        \item \(\overline{\zo} = z\) 
        \item \(|\Re z| \leq |z|\) and \(|\Im z|= \leq |z|\)
        \item \(|\zo| = |z|\) 
        \item \(|wz| = |w| |z|\)
    \end{enumerate}
\end{problem}

\begin{proof}
Suppose \(z = a + bi, w = c + di\). We verify these as follows:
\begin{enumerate}[label=(\alph*)]
    \item \(z + \zo = (a+bi) + (a-bi) = 2a = 2\Re z\)
    \item \(z - \zo = (a+bi) - (a-bi) = 2bi = 2 (\Im z)i\)
    \item \(z\zo = (a+bi)(a-bi)= a^2 + b^2 = |z|^2\)
    \item \(\overline{w+z} = \overline{(a+c)+(b+d)i} = (a+c) - (b+d)i = (a-bi)+(c-di) = \zo + \wo\)
    \item \(\overline{\zo}=\overline{a-bi}=a+bi = z\)
    \item \(|\Re z| = a^2 \leq a^2 + b^2 = |z|^2;\ \ |\Im z| = b^2 \leq a^2 + b^2 = |z|^2\)
    \item \(|\zo| = |a-bi| = |a+bi| = |z|\) 
    \item \(|wz| = |(c+di)(a+bi)| = | (ca - db) + (cb + da)i| = (ca - db)^2 + (cd + ba)^2\)
    This further equals that  \(|wz| = c^2a^2 +d^2b^2 + c^2d^2 + b^2a^2\).
    At the same time \(|w||z| = (c^2 + d^2)(a^2 + b^2) = c^2 a^2 + c^2 b^2 + d^2 a^2 + d^2 b^2\).
    Therefore they equal each other.
\end{enumerate}
\end{proof}

\begin{problem}{2}
    Prove the \textbf{reverse triangle inequality}: if \(w, z \in \C\), then 
    \(||w| - |z| | \leq |w - z|\). 
\end{problem}

\begin{proof}
Suppose \(z = a + b i, w = c + d i\). We have that 
\begin{align*}
    |w - z|^2 
    &= (w - z) (\wo - \zo) \\ 
    &= w\wo + z\zo - w\zo - z\wo \\ 
    &= |w|^2 + |z|^2 - w\zo - \overline{w \zo} \\ 
    &= |w|^2 + |z|^2 - 2\Re(w\zo) \\ 
    &\leq |w|^2 + |z|^2 - 2|w\zo| \\ 
    &= |w|^2 + |z|^2 - 2|w||z| \\ 
    &= (|w| - |z|)^2
\end{align*}
Taking the square root gets the desired equality. 
\end{proof}


\begin{problem}{3}
    Suppose \(V\) is a complex vector space and \(\varphi \in V'\). Define 
    \(\sigma \colon V \to \R\) by \(\sigma(v) = \Re \varphi(v)\) for each \(v \in V\). Show that 
    \[\varphi(v)  =\sigma(v) - i \sigma(iv)\]
    for all \(v \in V\). 
\end{problem}

\begin{proof}
\begin{align*}
    \varphi(v) 
    &= \Re \varphi(v) + \Im \varphi(v)  \\ 
    &= \sigma(v) + i \Re(- ii v) \\ 
    &= \sigma(v) - i \sigma(iv)
\end{align*}
\end{proof}


\begin{problem}{6}
    Suppose that \(m\) and \(n\) are positive integers wih \(m \leq n\), and suppose 
    \(\lambda_1, \ldots, \lambda_m \in \F\). Prove that there exists a polynomial 
    \(p \in \Pc(\F)\) with \(\deg p = n\) such that \(0 = p(\lambda_1) = \cdots = p(\lambda_m)\)
    and such that \(p\) has no other zeros. 
\end{problem}

\begin{proof}
We can first construct \(p(x) = (x - \lambda_1)(x - \lambda_2)\cdots(x-\lambda_m)\) for \(x \in \F\). 
This is a polynomial with degree \(m\) and has exactly \(m\) zeros. To increase the degree without 
introducing additional roots, we can make that \(p(x) = (x - \lambda_1)\cdots(x - \lambda_m)^{n-m+1}\). 
This polynomial doesn't have additional roots, as the proof procedure follows from the second 
version of fundamental theorem of algebra. 
\end{proof}



\begin{problem}{7}
    Suppose that \(m\) is a nonnegative integer, \(z_1, \ldots, z_{m+1}\) are distinct elements 
    of \(\F\), and \(w_1, \ldots, w_{m+1} \in \F\). Prove that there exists a unique polynomial 
    \(p \in \Pc_m(\F)\) such that 
    \[p(z_k) = w_k\]
    for each \(k = 1, \ldots, m+1\). 
\end{problem}

\begin{proof}
Define a linear map \(T \colon \Pc_m(\F) \to \F^{m+1}\) by 
\[T(p) = (p(z_1), \ldots, p(z_{m+1}))\]
We can tell that \(T\) is injective because if there exists nonzero \(p \in \nul T\), then 
this means there are \(m+1\) roots for the \(m\)-deg polynomial, forming a contradiction. Therefore, 
\(T\) is an isomorphism as the dimension matches. 

This implies the existence of the unique solution as the question asks. 
\end{proof}

\begin{problem}{9}
    Prove that every polynomial of odd degree with real coefficients has a real zero.
\end{problem}

\begin{proof}
Suppose for the sake of contradiction that there is no real root. Then the factorization would become 
\[p(x) = (x^2 + b_1x + c_1) \cdots (x^2 + b_Mx + c_M)\]
but this contradicts that the degree is odd. 
\end{proof}

\begin{problem}{11}
    Suppose \(p \in \Pc(\C)\). Define \(q \colon \C \to \C\) by 
    \[q(z) = p(z) \overline{p(\zo)}\]
    Prove that \(q\) is a polynomial with real coefficients. 
\end{problem}

\begin{proof}
Suppose \(p(z) = \sum_{i=0}^{n} a_i z^i\) for \(a_i \in \C, z \in \C\). Then we can find that 
\[p(\zo) = \sum_{i=0}^{n}a_i \zo^i\]
and thus  
\[\overline{p(\zo)} = \overline{\sum_{i=0}^{n}a_i \zo^i}
= \sum_{i=0}^{n} \ao_i z^i\]
Finally we can get that 
\[q(z) = p(z)\overline{p(\zo)} 
= (\sum_{i=0}^{n}a_i z^i)(\sum_{i=0}^{n}\ao_i z^i)\]

We can consider the general form of a coefficient of \(z^k\):
\[c_k = \sum_{i+j=k} a_i \ao_j = \sum_{j+i=k} \ao_j a_i = \overline{c_k}\]

Thus the coefficients are real.


\end{proof}





\end{document}

